A note on multi agent learning

### 2016-06-30 multiAgentLearning/paper/article.tex
added more text

\abstract{
  %TODO abstract
  \todo[inline]{abstract missing}
}
\todo{be more critical? maybe it depends to much on its paremeters?
  Or which game types are good for saticifing play?
}
\todo{misses callouts, intermezzi and quotes}
\section{Introduction}
\label{sec:introduction}
\toReview
Satisficing play was originally introduced
to try and limit the computation power that earlier models required of human
behavior \citep{herbert}. Originating from the field of economics, they
wanted to use a model that was not ``super intelligent''.

\toReview The classical models from the field of economics had at least one of the
following assumptions:
\begin{itemize}
\itemsep0em 
  \item The structure of the game is known.
  \item Awareness of other player's decisions.
  \item Awareness of being in a game situation
  \item Being able to preform maximization calculations.
\end{itemize}

\toReview These assumptions can be called common knowledge, and are a demanding to
make \citep{binmore}.

\todo{give paper overview}

@@

\toReview In the first phase the next action is selected. If the action still holds
according to the satisficing level or ambition it will remain the same, if not another
action will be selected. In the second phase the satisficing level is updated.
In here the constant learning rate is used. This is a variable that can be
modified to adjust the amount of new information being used or how much
information from memory is used.

\section{Satisficing play in action}
\toReview
To demonstrate a potential use of this algorithm we will use at several example
cases. The first one being the iterated prisoner's dilemma.

@@

\section{A saticificing comparison}
\drafting So, until now we've seen thoroughly how saticificing play works, and even some
strengths and weaknesses. However, what we really want to see is how it does
compare against other MAL algorithms. So to do this we'll first introduce the
competitors (Q-learning and WOLF-PHC) and then we'll show per game how they
do in comparison with help of the work of \citep{crandall}.

@@

The actual comparison of the algorithms is done by \citep{crandall}, in here
we will discuss several of his results. All the algorithms are compared in self
play. The algorithms use the following parameters:

\begin{tabular}{ll}
  Satisficing & $\alpha_0\in \mathbb{P}_0, \lambda=0.99$ \\
  WoLF-PHC & $\lambda=1/(10+0.001 k^a_s), \gamma = 0.95, \epsilon-\text{greedy w}$ 
             $/\epsilon = \tex{max}(0.2-0.0006t,0)$ \\
\end{tabular}

The first game result we'll discuss
is the already introduced prisoner's dillema:
\includegraphics[width=2.5in]{crandall/prisoners}

@@

In here we can see saticifing play performing a lot worse than the other two
algorithms. This is because, Crandall explains, the joint aspiration never
falls below the Pareto frontier, thus the behavior is essentially random.

\noindent
\fbox{
  \begin{minipage}{2.5in}
    \drafting
    \subsection{Multi-action saticificing}
    In their paper \citep{stimpson:2003} introduce a modification to
    saticificing play.
    Rather than just picking the other action, which it can't do because there
    are more than that available, it will pick uniformly across the other
    actions. This change allows saticificing play to play in $n$-action games
    rather than just 2 action games.
  \end{minipage}
}

@@

In here we can see saticificing play outperforming Q-Learning and WoLF quite
starkly. But we can also see that this result looks quite a lot like the
prisoners dillema result. This is because the MASD was basically a generalization
of the prisoners dillema, so this result isn't surprising.

\section{Properties of satisficing play}
To analyse saticfing play further we use the work of \citep{karandikar} and
\citep{stimpson:2001}. Karandikar gave some good theoretical foundations,
whereas stimpson did some deep emepircal experiments on self play in the
prisoners dillema with parameter perturbations.
\drafting

\todo{
Using both \citep{karandikar} and \citep{stimpson:2001} works to look deeper
at the properties of satisficing play.

karandikar used a theoretical proofing method, whereas Stimpson showed some
experimental results.

also showing the contrast graph which supports the work of \citep{arthur}
}

>>>>

\abstract{
  This paper takes a closer look at satisficing play. We explain the algorithm
  thoroughly and then look at empirical comparisons with Q-Learning and WoLF-PHC. From
  this we can conclude that satisficing play is relatively
  good at games with a cooperation aspect.
  Then we take a closer look at it in the prisoner's dillema. Where we
  discover that even with variance in the paremeters satisficing play ends up
  cooperating most of the time in self play. We conclude that satisficing play
  is an algorithm of serious interest to MAL researches.
}
\todo{be more critical? maybe it depends to much on its paremeters?
  Or which game types are good for satisficing play?
}
\todo{misses callouts and quotes}
\section{Introduction}
\label{sec:introduction}
\toReview
Satisficing play was originally introduced
to try and limit the computation power that earlier models required of human
behavior \citep{herbert}. Originating from the field of economics, they
wanted to use a model that was not ``super intelligent''.

The classical models from the field of economics had at least one of the
following assumptions:
\begin{itemize}
\itemsep0em 
  \item The structure of the game is known.
  \item Awareness of other player's decisions.
  \item Awareness of being in a game situation
  \item Being able to preform maximization calculations.
\end{itemize}

These assumptions can be called common knowledge, and are a demanding to
make \citep{binmore}.

In this paper we'll take a closer look at satisficing play. First we'll introduce
the core algorithm. Then we see how it works with several examples, which will
give us a better intuition about the algorithm. After that we see how it
compares to more ``classical'' MAL algorithms. Finally we dive deep into the
theoretical properties of satisficing play, and look at some empirical results
of satisficing play in the prisoner's dillema.

@@

In the first phase the next action is selected. If the action still holds
according to the satisficing level or ambition it will remain the same, if not another
action will be selected. In the second phase the satisficing level is updated.
In here the constant learning rate is used. This is a variable that can be
modified to adjust the amount of new information being used or how much
information from memory is used.

\section{Satisficing play in action}
\label{sec:playaction}
\toReview
To demonstrate a potential use of this algorithm we will use at several example
cases. The first one being the iterated prisoner's dilemma.

@@

\section{A saticificing comparison}
\toReview So, until now we've seen thoroughly how saticificing play works, and even some
strengths and weaknesses. However, what we really want to see is how it does
compare against other MAL algorithms. So to do this we'll first introduce the
competitors (Q-learning and WOLF-PHC) and then we'll show per game how they
do in comparison with help of the work of \citep{crandall}.

@@

The actual comparison of the algorithms is done by \citep{crandall}, in here
we will discuss several of his results. All the algorithms are compared in self
play. 
The first game result we'll discuss
is the already introduced prisoner's dillema:
\includegraphics[width=2.5in]{crandall/prisoners}

@@

In here we can see satisficing play performing a lot worse than the other two
algorithms. This is because, Crandall explains, the joint aspiration never
falls below the Pareto frontier, thus the behavior is essentially random.

\noindent
\fbox{
  \begin{minipage}{2.5in}
    \toReview
    \subsection{$N$-action saticificing}
    In their paper \citep{stimpson:2003} introduce a modification to
    saticificing play.
    Rather than just picking the other action, which it can't do because there
    are more than that available, it will pick uniformly across the other
    actions. This change allows saticificing play to play in $n$-action games
    rather than just 2 action games.
  \end{minipage}
}

@@

In here we can see saticificing play outperforming Q-Learning and WoLF quite
starkly. But we can also see that this result looks quite a lot like the
prisoner's dillema result. This is because the MASD was basically a generalization
of the prisoner's dillema, so this result isn't surprising.

\section{Properties of satisficing play}
\toReview
To analyse saticfing play further we use the work of \citep{karandikar} and
\citep{stimpson:2001}. Karandikar gave some good theoretical foundations,
whereas stimpson did empirical experiments on self play in the
prisoner's dillema with parameter perturbations. However as said in the
\nameref{sec:core-algorithm} it should be noted
that the algorithms of Karandikar and Stimpson are different. The main
difference is the lack of \emph{inertia} in the algorithm of Stimpson, which
makes the behavior deterministic.

Karandikar sums up his work quite thoroughly in  the informal discussion section
of his paper. First of all ``   First, the untrembled joint process of
aspirations and actions always converges to a pure strategy state.''
This means that even without trembles, saticfing play with inertia
will end up in a state wherein aspiration levels of both players equals payoff.
The second result I want to discuss is:
``The limit invariant distribution places
almost all weight on the cooperative outcome, provided that the persistence
parameter is sufficiently close to unity.''. Where the persistence parameter
is the learning rate, and the invariant distribution is the trembled process,
in other words, the saticfing play algorithm with trembles. For more details
the reader is referred towards that paper.

Before the results of Stimpson are discussed we first have to consider the
possible outcomes. It is important to keep in mind that in his algorithm
the result of karandikar doesn't hold, because inertia is missing. Therefore
not all processes will converge to a pure strategy state.
the possible outcomes are:
\begin{enumerate}
  \item Convergence to certain actions.
  \item Action cycles.
  \item Complete chaos.
\end{enumerate}
We've seen both one and two in the section \nameref{sec:playaction}. Three
is mentioned because there lacks a theoretical proof of it being impossible.
Stimpson studied the behavior of saticfing play empirically in the prisoner's
dillema, so over 5000 trials he ended up with the following result:

\includegraphics[width=2.5in]{stimpson/strategies}

Where the leters stand for the actions, for example DC is (Defect,Cooperate).
If letters are have dashes in between they indicate action cycles, for example:
$DD-CC-DC$ is (Defect, Defect), (Cooperate, Cooperate) and 
(Defect,Cooperate).
From here we can see that saticfing play ends up cooperating most of the time.
The variations are because the parameters were set randomly, such as initial
aspiration, learning rate or even initial action.

They also made another graph depicting how initial aspiration levels influenced
the perecentage of the time ending up with mutual cooperation, this happened
over 1000 trials:

\includegraphics[width=2.5in]{stimpson/aspiration}

Alfa is the initial aspiration of player 1, Beta of player 2, and the colors
indicate the percentage of the time that cooperation happened. From this we can
see that if the goal is cooperation, having initially high aspiration is good
most of the time. You can also see some kind of intrusion towards the center,
from which you could conclude that having similiar aspirations is good for
cooperation. But the authors do not do this. It is also somewhat crenelated,
probably because of the low number of trials they used to make this graph.

To further study satisficing play in the prisoner's dillema, a generalization
has to be made. Therefore stimpson introduces a generalized dillema that looks
like the following normal form game:
\begin{tabular}{lll}
            & Cooperate   & Defect \\
  Cooperate & $(\sigma,\sigma)$     & $(0,1)$  \\
  Defect    & $(1,0)$     & $(\delta,\delta)$  \\
\end{tabular}
Where $(0<\delta<\sigma<1)$ and $\sigma>0.5$. Sigma $\sigma$ for cooperate and
Delta $\delta$ for defect. The constraint $(0<\delta<\sigma<1)$ exists to make
sure the game is a prisoner's dillema and $\sigma>0.5$ is used to ensure the
Pareto front of the prisoner's dillema.
The following graph illustrates the effect of different reward contrasts,
and was created over 1000 trials:

\noindent
\includegraphics[width=2.5in]{stimpson/contrast}

The colors indicate the percentage of time a run resulted in cooperation.
According to the authors, cooperation is most likely when $\delta$ is small
and $\sigma$ is large. If they are close together distinction between 
defection and cooperation becomes difficult. This is typical of non-optimizing
algorithms, and Arthur writes that his algorithm and even human behavior

\multicolinterrupt{
\begin{center}
\large\emph{``Beyond the perceptual threshold, non-optimal outcomes become likely''}
\end{center}
\vspace{-2em}\hbox{}\hfill --- \citep{arthur}
}

``appear to discover and exploit the optimal action with high probability,
\emph{as long as it is not difficult to discriminate}. But beyond a perceptual
threshold, where differences in alternatives become less pronounced,
non-optimal outcomes become more likely'' \citep{arthur}

The final result presented by stimpson is the learning rate. The learning
rate determines how quickly aspiration should be updated. A higher learning
rate means that values of the past are less relevant. The following graph
was made over 1000 trials:

\noindent
\includegraphics[width=2.5in]{stimpson/learning-rate}

Where lambda is the learning rate. As we can see the percentage of cooperation
(the number on the $y$-axis), increases with higher learning rate. But starkly
drops once it reaches one and aspiration isn't updated at all.

\section{Conclusion}
\toReview

We've seen how saticfing play works and some potential issues with it.
Then we compared it against other multi-agent
algorithm's empircally. Where we've
seen it be better in a lot of famous games, but in others such as shapely it
performed relatively more poorly. We've also seen it been extended to play
with more actions and against more agents in the MASD.
After that we looked closer to the theoretical properties
of saticfing play, where we've seen it with the introduction of random elements
convergence to the coopertive outcome becomes likely.
Finally we've also seen empirical tests
on the deterministic variant in the prisoner's dillema. From this we've seen
that parameters are important, but most of the time it results in the coopertive
outcome anyway.

From this we can conclude that saticfing play is a serious algorithm to be of
interest of MAL researches.
